Cluster Difference Imaging Photometric Survey. IV. TOI 1937Ab. Youngest Hot Jupiter, or Just Tidally Spun Up?

### 2021-03-01 cluster_paper/ms.tex
update reddening

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$120\,Myr open cluster NGC\,2516.
  %TODO: says who?
  The classical tidal radius of the cluster is $\approx$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf NGC\,2516 has multiple reported components in position and
  velocity space.} The core (black) was identified by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident with
  the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo was identified by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021} using less restrictive membership
  assignment algorithms (see Appendix~\ref{app:clustering}).  The field
  sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
  centered on the cluster.  The low volume density of the
  halo stars makes it difficult to visually distinguish
  the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: The photometric data are from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars having not yet reached the
    main-sequence.
    {\it Right}: Solar-metallicity PARSEC isochrones are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$ with a
    reddening correction applied.
    The main-sequence turn-off location and evolved stars suggests a
    cluster age slightly older than the Pleiades.  The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}} to interpolate these isochrones
at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and then applied the
reddening correction described in Section~\ref{subsec:tess}, in the
sense of adding the $E(Bp-Rp)$ correction to the synthetic Gaia EDR3
photometry.  The data and models generally agree for masses above
approximately 0.45\,$M_\odot$.  Below this mass, at colors redder than
Bp-Rp$\approx2.2$, the agreement is worse.  The MIST isochrones show a
comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
both starspots and incomplete molecular line lists have been proposed
\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
For purposes of age-dating the cluster, we prefer to simply focus on
the main-sequence turn-off (MSTO), since it is where the models have maximal
predictive power.

Our photometric age-dating procedure is as follows.  We selected all
stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
outliers from the rest of the cluster.  Using an isochrone grid
spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
spacing, we then numerically evaluated the $\chi^2$ fit for each
possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
percentiles are used as summary statstics.

% NOTE: probably asymmetric.
The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.

@@

		% \subfloat{
		% 	\includegraphics[width=0.75\textwidth]{f3b.pdf}
		% }
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    % The {\it top} plot shows periods against a linear scale, while the
    % {\it bottom}  shows them against a logarithmic scale.
    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
    and Praesepe \citep[650\,Myr][]{douglas_poking_2017}
    are shown for reference.
    Stars in the core and the halo of NGC\,2516 are
    gyrochronologically coeval with the Pleiades.
		\label{fig:rot}
	}
\end{figure*}

>>>>

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$175\,Myr open cluster NGC\,2516.
  %TODO: says who?
  The classical tidal radius of the cluster is $\approx$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across position and
  velocity space.} The core (black) was identified by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident with
  the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) was identified by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021} using less restrictive membership
  assignment algorithms described in Appendix~\ref{app:clustering}.  The field
  sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
  centered on the cluster.  The low volume density of the
  halo stars makes it difficult to visually distinguish
  the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: The photometric data are from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars having not yet reached the
    main-sequence.
    {\it Right}: Solar-metallicity PARSEC isochrones are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
    The main-sequence turn-off location and evolved stars suggests a
    cluster age slightly older than the Pleiades.  The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}} to interpolate these isochrones
at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied the
reddening correction described in Section~\ref{subsec:tess}. This
entailed by including using our corrected value of
$E(B-V)=0.102\pm0.019$, assuming $R_V=3.1$, and applying the
\citet{odonnell_1994} SED-dependent reddening law.
The data and models generally agree for masses above
approximately 0.45\,$M_\odot$.  Below this mass, at colors redder than
Bp-Rp$\approx2.2$, the data and models diverge.  The MIST isochrones show a
comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
both starspots and incomplete molecular line lists have been proposed
\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the 
PARSEC isochrones since they diverge at slightly lower mass than the
MIST models, due to the temperature-opacity calibration
discussed by \citet{chen_improving_2014}.
For purposes of age-dating the cluster, we prefer to simply focus on
the main-sequence turn-off (MSTO), since it is where the models have
maximal predictive power.

\citet{cummings_2018} presented techniques for mitigating some of the
complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
rotation, high binarity fractions, and the presence of blue
stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
cluster memberships, they found MSTO ages for NGC\,2516 ranging from
165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
PARSEC, MIST, or SYCLIST).  The statistical uncertainties ($\approx
25$\,Myr) were comparable to the systematic uncertainties.

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} shows
empirically that they seem to be: stars past the turnoff in both the
halo and core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} suggests the precision with which the claim can be
made.  The MSTO stars are consistent with the 178 and 316 Myr models,
but are bluer than the 100 Myr model.  The RGB stars are consistent
only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$) model.  Based on
the assumption that the width of the MSTO can be attributed to binary
stars, we are most interested in the blue edge. This seems most
compatible with the 178 Myr model.

% Our photometric age-dating procedure is as follows.  We selected all
% stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
% excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
% outliers from the rest of the cluster.  Using an isochrone grid
% spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
% spacing, we then numerically evaluated the $\chi^2$ fit for each
% possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
% percentiles are used as summary statstics.
% 
% % NOTE: probably asymmetric.
% The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
% For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
% yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
% the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.

@@

		% \subfloat{
		% 	\includegraphics[width=0.75\textwidth]{f3b.pdf}
		% }
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    % The {\it top} plot shows periods against a linear scale, while the
    % {\it bottom}  shows them against a logarithmic scale.
    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a}
    and Praesepe \citep[650\,Myr][]{douglas_poking_2017}
    are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
    in which some NGC\,2516 members appear elevated with respect to
    the Pleiades, suggesting a slightly older age.
		\label{fig:rot}
	}
\end{figure*}

### 2021-03-01 planet_paper/ms.tex
update reddening

Processes that have been suggested to produce hot Jupiters occur on
timescales ranging from megayears to gigayears (CITE, CITE, CITE).
One way to distinguish between these processes is therefore to find
young ($\lesssim$100\,Myr) hot Jupiters.  However the search for
sub-100\,Myr hot Jupiters has yielded remarkably few, if any, secure
detections.

Transits have provided a pile of tantalizing candidates.  HIP\,67522b,
with an age, orbital period, and size of XX\,Myr, X.X\,days, and
TOOSMALL\,$R_\oplus$ respectively, comes close
\citep{rizzuto_tess_2020}.  Its
size however is smaller than expected for a Jovian-mass object given
its age, by a factor of $\approx1.5\times$ (CITE Burrows, Fortney,
Thorngren models, CITE Owen 2020 entropy).  A mass measurement is
therefore needed to resolve its status as either a hot Jupiter, or an
inflated Neptunian world.  No other planet comes as close to fitting
the bill.  V1298 Tau b, part of a unique system with at least four
transiting planets, has the correct size and age. However its orbital
period (XX\,days) is too long to be a hot Jupiter (CITE David 2019,
2020).  The dips in PTFO 8-8695, long-interpreted as a candidate hot
Jupiter, are chromatic and show a changing orbital phase.  This is
inconsistent with a planetary interpretation (van Eyken 2012, Yu+15,
Onitsuka+17, Tanimoto+20, Bouma+20).

Reports of very young hot Jupiters have come from the radial velocity
(RV) technique, but these reports are difficult to verify.  TW Hya
(age=X\,Myr), for instance, showed a radial velocity semi-amplitude of
$\approx$200\,m\,s$^{-1}$ with a period of 3.6\,days, as measured in
optical spectra \citep{setiawan_young_2008} (CITE age reference).
Even though \citet{setiawan_young_2008} observed no significant
correlation between activity indicators and the RV signal, subsequent
infrared velocities measured by \citet{huelamo_tw_2008} showed small
variations $\lesssim 35\,$m\,s$^{-1}$.  \citet{huelamo_tw_2008}
ultimately concluded that activity-induced variations were the source
of the optical signal.  More recently, very young hot Jupiters have
been reported around CI Tau, V830 Tau, and TAP 26
\citep{johns-krull_CI_Tau_candidate_2016,donati_hot_2016,donati_hot_2017,yu_hot_2017,biddle_k2_2018,flagg_co_2019}.
The planetary nature of at least two of these signals (CI Tau and V830
Tau) has been debated \citep{donati_magnetic_2020,damasso_gaps_2020}.
More generally, the Bayesian model comparison problem of ``stellar
activity alone'' versus ``stellar activity plus planet'' has been
shown to require significant amounts of both data and statistical
machinery \citep[{\it
e.g.},][]{barragan_radial_2019,klein_simulated_2020} To date, none of
the reported young hot Jupiter detections have performed this style of
analysis.  Such an analysis, or else the acquisition of multi-color
radial velocities, seems like a requirement given the challenges of
detecting small signals in the presence of significant stellar
variability.

So where does that leave the search for young hot Jupiters?
A quick query to the NASA Exoplanet Archive yields the current state
of the search.
(FIGURE: Rp vs Age, and Mp\,sini vs Age).

>>>>

Depending on the process that produces them, hot Jupiters arrive on
their tiny orbits on timescales of anywhere between megayears and
gigayears (CITE, CITE, CITE).
One way to distinguish between these processes is to find
young ($\lesssim$100\,Myr) hot Jupiters.
However the search for such hot Jupiters has yielded remarkably few,
if any, secure detections.
What is the youngest hot Jupiter known?

Transits have provided a pile of tantalizing candidates.  HIP\,67522b,
with an age, orbital period, and size of XX\,Myr, X.X\,days, and
TOOSMALL\,$R_\oplus$ respectively, comes close
\citep{rizzuto_tess_2020}.  Its
size however is smaller than expected for a Jovian-mass object given
its age, by a factor of $\approx1.5\times$ (CITE Burrows, Fortney,
Thorngren models, CITE Owen 2020 entropy).  A mass measurement is
therefore needed to resolve its status as either a hot Jupiter, or an
inflated Neptunian world.  No other planet comes as close to fitting
the bill.  V1298 Tau b, part of a unique system with at least four
transiting planets, is appropriately large and young. However its orbital
period (XX\,days) is too long to be a hot Jupiter (CITE David 2019,
2020).  The dips in PTFO 8-8695, long-interpreted as a candidate hot
Jupiter, are chromatic and show a changing orbital phase.  This is
inconsistent with a planetary interpretation (van Eyken 2012, Yu+15,
Onitsuka+17, Tanimoto+20, Bouma+20).

Reports of very young hot Jupiters have come from the radial velocity
(RV) technique, but these reports are difficult to verify.  TW Hya
(age=X\,Myr), for instance, showed a radial velocity semi-amplitude of
$\approx$200\,m\,s$^{-1}$ with a period of 3.6\,days, as measured in
optical spectra \citep{setiawan_young_2008} (CITEP ALSO age reference).
Even though \citet{setiawan_young_2008} observed no significant
correlation between activity indicators and the RV signal, subsequent
infrared velocities measured by \citet{huelamo_tw_2008} showed small
variations $\lesssim 35\,$m\,s$^{-1}$.  \citet{huelamo_tw_2008}
ultimately concluded that activity-induced variations were the source
of the optical signal.  More recently, very young hot Jupiters have
been reported around CI Tau, V830 Tau, and TAP 26
\citep{johns-krull_CI_Tau_candidate_2016,donati_hot_2016,donati_hot_2017,yu_hot_2017,biddle_k2_2018,flagg_co_2019}.
The planetary nature of at least two of these signals (CI Tau and V830
Tau) has been debated \citep{donati_magnetic_2020,damasso_gaps_2020}.
More generally, the Bayesian model comparison problem of ``stellar
activity alone'' versus ``stellar activity plus planet'' has been
shown to require significant amounts of both data and statistical
machinery \citep[{\it
e.g.},][]{barragan_radial_2019,klein_simulated_2020} To date, none of
the reported young hot Jupiter detections have performed this style of
analysis.  Such an analysis, or else the acquisition of multi-color
radial velocities, seems like a requirement given the challenges of
detecting small signals in the presence of significant stellar
variability.

%TODO: make figure
So where does that leave the search for young hot Jupiters?
A quick query to the NASA Exoplanet Archive yields the current state
of the search, showcased in Figure~\ref{fig:rp_vs_age}.
(FIGURE: Rp vs Age, and Mp\,sini vs Age).
The youngest transiting hot Jupiter ($R_p>R_{\rm Jup}$, $P<10\,{\rm
days}$) appears to be Qatar-4b. 
Discovered by (CITE), the age was reported to be very low because of
gyrochrones.
Not great (CITE CITE).
